Modelling BCG vaccination in the UK

### 2019-07-04 rmds/04-tb-epi-england.Rmd
shortened table and figure captions in lists

```{r data-struct, fig.cap = "Summary plot of the structure of the ETS data extract used in this thesis. Due to the large size of the dataset, the data has been sub sampled with only 20\\% of the data shown. Notifications have been ordered by date of notification from left to right. Purple indicates numeric variables, green indicates factor variables, dates are shown using red and samples with missing values are shown using grey. The figure shows that the majority of the included variables are factors, with a significant minority of numeric and date variables. Data completeness is near random for some variables but has changed other time for others, for example national socioeconomic quintile (natquintile). This figure does not account for nested variables", out.width = "60%", eval = FALSE}
if (regen_results) {
  plot_data_struct <- ets %>% 
    sample_frac(0.2) %>% 
    arrange(desc(caserepdate)) %>% 
  vis_dat(warn_large_data = FALSE) + 
    coord_flip() +
    theme(legend.position = "bottom")

@@

```{r plot-missing-struct, fig.cap = "Summary plot of missing data in the extract of the ETS data used in this thesis. Due to the large size of the dataset, the data has been sub-sampled with only 20\\% of the data shown in this figure. Notifications have been ordered by date of notification from left to right. The following subset of variables are shown; year (year), sex (sex), age (age), PHE Centre (phec), Occupation (occat), Ethnic group (ethgrp), UK birth status (ukborn), Time since entry (timesinceent), date of symptom onset (symptonset), date of diagnosis (datediag), started treatment (startedtreat), date of starting treatment (starttreatdate), treatment end date (txenddate), pulmonary or extra-pulmonary TB (pulmextrapulm), culture (culture), sputum smear status (sputsmear), drug resistance (anyres), previous diagnosis (prevdiag), BCG status(bcgvacc), Year of BCG vaccination (bcgvaccyr), overall outcome (overalloutcome), cause of death (tomdeathrelate), socio-economic status quintiles (natquintile), and date of death (dateofdeath). Nested variables have been accounted for (i.e data of death has had an entry added for cases that are known to have not died), so that true missingness for all variables is estimated.", out.width = "60%"}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-missing-struct.png")
```

@@

kable(miss_ets,
      caption = "Breakdown of missing data from the ETS prior to the web based system (pre 2009) and post (post 2008) by variable, ordered by the  percentage missing for a subset of variables. The following subset of variables are shown; year (year), sex (sex), age (age), PHE Centre (phec), Occupation (occat), Ethnic group (ethgrp), UK birth status (ukborn), Time since entry (timesinceent), date of symptom onset (symptonset), date of diagnosis (datediag), started treatment (startedtreat), date of starting treatment (starttreatdate), treatment end date (txenddate), pulmonary or extra-pulmonary TB (pulmextrapulm), culture (culture), sputum smear status (sputsmear), drug resistance (anyres), previous diagnosis (prevdiag), BCG status(bcgvacc), Year of BCG vaccination (bcgvaccyr), overall outcome (overalloutcome), cause of death (tomdeathrelate), socio-economic status quintiles (natquintile), and date of death (dateofdeath). Nested variables have been accounted for (i.e data of death has had an entry added for cases that are known to have not died), so that true missingness for all variables is estimated.",
      booktabs = TRUE) %>%
  add_header_above(c(" ", "Pre 2009" = 2, "Post 2008" = 2)) %>% 
  kable_styling(latex_options = "HOLD_position") 
```

@@

```{r bcgvacc-miss}
kable(model_miss_results[["bcgvacc"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that BCG status is missing at random for the variables considered.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, 
                latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r bcgvaccyr-miss}
kable(model_miss_results[["bvcvaccyr"]], format = "latex",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for year of BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that year of BCG vaccination is missing at random for the variables considered.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm")
```

@@

```{r dateofdeath-miss}
kable(model_miss_results[["dateofdeath"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that there is some evidence that date of death is missing at random for ethnic group, with weaker evidence for all other variables.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r tomdeathrealat-miss}
kable(model_miss_results[["tomdeathrealat"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for cause of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that cause of death is missing at random for ethnic group and UK birth status, with little evidence for any other variables", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r symptonset-miss}
kable(model_miss_results[["symptonset"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of symptom onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that date of symptom onset is missing not at random for the variables for all variables considered, except for sex.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r datediag-miss}
kable(model_miss_results[["datediag"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of diagnosis onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that date of diagnosis is missing at random for the variables for all variables considered, except for sex.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r starttreatdate-miss}
kable(model_miss_results[["starttreatdate"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. There is little evidence that the missing data for the date of starting treatment is associated with any variable considered, except for year of notification.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r txenddate-miss}
kable(model_miss_results[["txenddate"]],
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. There is little evidence that the missing data for the date of starting treatment is associated with any variable considered, except for year of notification.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r plot-detection-not, fig.cap = "a.) and b.) show notifications over time by date of notification in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases notified in a given month for each year, with some evidence of a seasonal trend. d.) Shows the proportion of cases notified on a given day for each month, there is little evidence of between day variation in cases notified."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-not.png")
```

@@

```{r plot-detection-sympton, fig.cap = "a.) and b.) show notifications over time by date of symptom onset in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases notified in a given month for each year, with some evidence of a seasonal trend and a higher proportion of cases reporting symptoms starting in January than would be expected. d.) Shows the proportion of cases notified on a given day for each month, with a much higher proportion of cases reproting symptoms on the first of the month than would be expected. On both the scale of months and years there is some evidence of recall bias, with the first month, or first day, reporting higher proportions of cases than would be expected."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-sympton.png")
```

@@

```{r plot-detection-starttreatdate, fig.cap = "a.) and b.) show notifications over time by date of starting treatment in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases starting treatment in a given month for each year, with some evidence of a seasonal trend. d.) Shows the proportion of cases starting treatment on a given day for each month, with little evidence of between day variation. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-starttreatdate.png")
```

@@

```{r plot-detection-treatend, fig.cap = "a.) and b.) show notifications over time by date of treatment ending in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases finishing treatment in a given month for each year, with little evidence of a seasonal trend. d.) Shows the proportion of cases finishing treatment on a given day for each month, with a much higher proportion of cases finishing treatment on the first of the month than would be expected. On the scale of months there is some evidence of recall bias, with the first day reporting higher proportions of cases than would be expected. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-treatend.png")
```

@@

```{r plot-detection-dateofdeath, fig.cap = "a.) and b.) show notifications over time by date of death in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases who died in a given month for each year, with no evidence of a seasonal trend. d.) Shows the proportion of cases who died on a given day for each month, with little evidence of between day variation. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-dateofdeath.png")
```

@@

```{r plot-overall-pop, fig.cap = "Overall population estimates derived using Office for National Statistics (ONS) and LFS demographic data. The ONS data is likely to be more reliable as the LFS data is derived using a weighted survey. After accounting for missing UK birth status both datasets provide comparable estimates of the population of England, with a clearly increasing trend over time. However the ONS data indicates a reduction in population from 2000 until 2001 that is not seen in the LFS data. The UK born population has also risen slowly from 2000 until 2015, although the biggest increase has been in the non-UK born population."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-pop.png")
```

```{r plot-age-strat, fig.cap = "Percentage difference between Office for National Statistics (ONS) population estimates and estimates derived from the LFS by 5 year age group. For most age groups there is less than a 2\\% difference over time. In older adults (85+) there is a substantially greater difference ranging from 5\\% to 40\\%."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-age-strat.png")
```

@@

```{r plot-overall-nots, fig.cap = "Notifications in England from 2000 to 2015 stratified by UK birth status, sourced from the ETS system. Notifications in the non-UK born doubled from 3329 in 2000 to 6021 in 2011 since when they have decreased year on year. In the UK born notifications have remained comparable over time, with some evidence of a decrease from 2011 until 2015. UK birth status has become increasingly complete over time with notifications without birth status dropping from 885 in 2000 to 121 in 2015."}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-nots.png")
```

@@

```{r age-dist-cases-nots, fig.cap = "Proportion of cases by 5 year age group in the ETS system in 2005, 2010 and 2015 stratified by UK birth status. Non-UK born cases have a higher proportion of young adult cases with very few cases in children or in older adults. UK born cases have a more uniform distribution of cases with some evidence of a higher proportion of cases in young adults. In the non-UK born the proportion of cases in young adults has decreased over time, with no evidence of a temporal trend in the UK born. These results are not adjusted for population demographics and therefore may be biased."}
if (regen_results) {

>>>>

```{r data-struct, fig.scap = "Summary plot of the structure of the ETS data extract used in this thesis.", fig.cap = "Summary plot of the structure of the ETS data extract used in this thesis. Due to the large size of the dataset, the data has been sub sampled with only 20\\% of the data shown. Notifications have been ordered by date of notification from left to right. Purple indicates numeric variables, green indicates factor variables, dates are shown using red and samples with missing values are shown using grey. The figure shows that the majority of the included variables are factors, with a significant minority of numeric and date variables. Data completeness is near random for some variables but has changed other time for others, for example national socioeconomic quintile (natquintile). This figure does not account for nested variables", out.width = "60%", eval = FALSE, out.extra = ""}
if (regen_results) {
  plot_data_struct <- ets %>% 
    sample_frac(0.2) %>% 
    arrange(desc(caserepdate)) %>% 
  vis_dat(warn_large_data = FALSE) + 
    coord_flip() +
    theme(legend.position = "bottom")

@@

```{r plot-missing-struct, fig.scap = "Summary plot of missing data in the extract of the ETS data used in this thesis.", fig.cap = "Summary plot of missing data in the extract of the ETS data used in this thesis. Due to the large size of the dataset, the data has been sub-sampled with only 20\\% of the data shown in this figure. Notifications have been ordered by date of notification from left to right. The following subset of variables are shown; year (year), sex (sex), age (age), PHE Centre (phec), Occupation (occat), Ethnic group (ethgrp), UK birth status (ukborn), Time since entry (timesinceent), date of symptom onset (symptonset), date of diagnosis (datediag), started treatment (startedtreat), date of starting treatment (starttreatdate), treatment end date (txenddate), pulmonary or extra-pulmonary TB (pulmextrapulm), culture (culture), sputum smear status (sputsmear), drug resistance (anyres), previous diagnosis (prevdiag), BCG status(bcgvacc), Year of BCG vaccination (bcgvaccyr), overall outcome (overalloutcome), cause of death (tomdeathrelate), socio-economic status quintiles (natquintile), and date of death (dateofdeath). Nested variables have been accounted for (i.e data of death has had an entry added for cases that are known to have not died), so that true missingness for all variables is estimated.", out.width = "60%", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-missing-struct.png")
```

@@

kable(miss_ets,
      caption.short = "Breakdown of missing data from the ETS prior to the web based system (pre 2009) and post (post 2008) by variable, ordered by the  percentage missing for a subset of variables.",
      caption = "Breakdown of missing data from the ETS prior to the web based system (pre 2009) and post (post 2008) by variable, ordered by the  percentage missing for a subset of variables. The following subset of variables are shown; year (year), sex (sex), age (age), PHE Centre (phec), Occupation (occat), Ethnic group (ethgrp), UK birth status (ukborn), Time since entry (timesinceent), date of symptom onset (symptonset), date of diagnosis (datediag), started treatment (startedtreat), date of starting treatment (starttreatdate), treatment end date (txenddate), pulmonary or extra-pulmonary TB (pulmextrapulm), culture (culture), sputum smear status (sputsmear), drug resistance (anyres), previous diagnosis (prevdiag), BCG status(bcgvacc), Year of BCG vaccination (bcgvaccyr), overall outcome (overalloutcome), cause of death (tomdeathrelate), socio-economic status quintiles (natquintile), and date of death (dateofdeath). Nested variables have been accounted for (i.e data of death has had an entry added for cases that are known to have not died), so that true missingness for all variables is estimated.",
      booktabs = TRUE) %>%
  add_header_above(c(" ", "Pre 2009" = 2, "Post 2008" = 2)) %>% 
  kable_styling(latex_options = "HOLD_position") 
```

@@

```{r bcgvacc-miss}
kable(model_miss_results[["bcgvacc"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that BCG status is missing at random for the variables considered.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, 
                latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r bcgvaccyr-miss}
kable(model_miss_results[["bvcvaccyr"]], format = "latex",
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for year of BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for year of BCG vaccination as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that year of BCG vaccination is missing at random for the variables considered.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm")
```

@@

```{r dateofdeath-miss}
kable(model_miss_results[["dateofdeath"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for date of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that there is some evidence that date of death is missing at random for ethnic group, with weaker evidence for all other variables.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r tomdeathrealat-miss}
kable(model_miss_results[["tomdeathrealat"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for cause of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for cause of death as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that cause of death is missing at random for ethnic group and UK birth status, with little evidence for any other variables", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r symptonset-miss}
kable(model_miss_results[["symptonset"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for date of symptom onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of symptom onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that date of symptom onset is missing not at random for the variables for all variables considered, except for sex.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r datediag-miss}
kable(model_miss_results[["datediag"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for date of diagnosis onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).", 
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of diagnosis onset as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. The model indicates that date of diagnosis is missing at random for the variables for all variables considered, except for sex.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r starttreatdate-miss}
kable(model_miss_results[["starttreatdate"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).", 
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. There is little evidence that the missing data for the date of starting treatment is associated with any variable considered, except for year of notification.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r txenddate-miss}
kable(model_miss_results[["txenddate"]],
      caption.short = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles).",
      caption = "Results from a logistic regression model with data completeness (Complete/Missing) for date of starting treatment as an outcome, adjusted for; year, sex, age (grouped as 0-14 year olds, 15-65 year olds and 65+), ethnic group, UK birth status and socio-economic status (national quintiles). For socio-economic group 1 indicates the most deprived quintile. Notifications from 2010 onwards were included as socio-economic status was not collected before this. Complete case analysis weas used. Odds ratios shown are adjusted for all explanatory variables. There is little evidence that the missing data for the date of starting treatment is associated with any variable considered, except for year of notification.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(font_size = 8, latex_options = c("repeat_header")) %>% 
  column_spec(c(4), width = "2cm") %>% 
  column_spec(c(6, 7), width = "1.5cm") %>% 
  kable_styling(latex_options = "hold_position")
```

@@

```{r plot-detection-not, fig.scap = "a.) and b.) show notifications over time by date of notification in the ETS, with a.) aggregated by year and  b.) aggregated by month.", fig.cap = "a.) and b.) show notifications over time by date of notification in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases notified in a given month for each year, with some evidence of a seasonal trend. d.) Shows the proportion of cases notified on a given day for each month, there is little evidence of between day variation in cases notified.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-not.png")
```

@@

```{r plot-detection-sympton, fig.scap = "a.) and b.) show notifications over time by date of symptom onset in the ETS, with a.) aggregated by year and  b.) aggregated by month.", fig.cap = "a.) and b.) show notifications over time by date of symptom onset in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases notified in a given month for each year, with some evidence of a seasonal trend and a higher proportion of cases reporting symptoms starting in January than would be expected. d.) Shows the proportion of cases notified on a given day for each month, with a much higher proportion of cases reproting symptoms on the first of the month than would be expected. On both the scale of months and years there is some evidence of recall bias, with the first month, or first day, reporting higher proportions of cases than would be expected.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-sympton.png")
```

@@

```{r plot-detection-starttreatdate, fig.scap = "a.) and b.) show notifications over time by date of starting treatment in the ETS, with a.) aggregated by year and  b.) aggregated by month.", fig.cap = "a.) and b.) show notifications over time by date of starting treatment in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases starting treatment in a given month for each year, with some evidence of a seasonal trend. d.) Shows the proportion of cases starting treatment on a given day for each month, with little evidence of between day variation. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-starttreatdate.png")
```

@@

```{r plot-detection-treatend, fig.scap = "a.) and b.) show notifications over time by date of treatment ending in the ETS, with a.) aggregated by year and  b.) aggregated by month.", fig.cap = "a.) and b.) show notifications over time by date of treatment ending in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases finishing treatment in a given month for each year, with little evidence of a seasonal trend. d.) Shows the proportion of cases finishing treatment on a given day for each month, with a much higher proportion of cases finishing treatment on the first of the month than would be expected. On the scale of months there is some evidence of recall bias, with the first day reporting higher proportions of cases than would be expected. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-treatend.png")
```

@@

```{r plot-detection-dateofdeath, fig.scap = "a.) and b.) show notifications over time by date of death in the ETS, with a.) aggregated by year and  b.) aggregated by month.", fig.cap = "a.) and b.) show notifications over time by date of death in the ETS, with a.) aggregated by year and  b.) aggregated by month. A trendline has been produced using a locally weighted regression model. Both of these plots show the same overall trend, but b.) contains a large amount of apparent noise. c.) Shows the proportion of cases who died in a given month for each year, with no evidence of a seasonal trend. d.) Shows the proportion of cases who died on a given day for each month, with little evidence of between day variation. Data is only shown from 2001 until 2015 and prior to 2001 this variable was not recorded and it is not complete for 2015.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-detection-dateofdeath.png")
```

@@

```{r plot-overall-pop, fig.scap = "Overall population estimates derived using ONS and LFS demographic data.", fig.cap = "Overall population estimates derived using ONS and LFS demographic data. The ONS data is likely to be more reliable as the LFS data is derived using a weighted survey. After accounting for missing UK birth status both datasets provide comparable estimates of the population of England, with a clearly increasing trend over time. However the ONS data indicates a reduction in population from 2000 until 2001 that is not seen in the LFS data. The UK born population has also risen slowly from 2000 until 2015, although the biggest increase has been in the non-UK born population.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-pop.png")
```

```{r plot-age-strat, fig.scap = "Percentage difference between ONS population estimates and estimates derived from the LFS by 5 year age group.", fig.cap = "Percentage difference between ONS population estimates and estimates derived from the LFS by 5 year age group. For most age groups there is less than a 2\\% difference over time. In older adults (85+) there is a substantially greater difference ranging from 5\\% to 40\\%.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-age-strat.png")
```

@@

```{r plot-overall-nots, fig.scap = "Notifications in England from 2000 to 2015 stratified by UK birth status, sourced from the ETS system.", fig.cap = "Notifications in England from 2000 to 2015 stratified by UK birth status, sourced from the ETS system. Notifications in the non-UK born doubled from 3329 in 2000 to 6021 in 2011 since when they have decreased year on year. In the UK born notifications have remained comparable over time, with some evidence of a decrease from 2011 until 2015. UK birth status has become increasingly complete over time with notifications without birth status dropping from 885 in 2000 to 121 in 2015.", out.extra = ""}
knitr::include_graphics("chapters/tb-epi-england/figures/plot-overall-nots.png")
```

@@

```{r age-dist-cases-nots, fig.scap = "Proportion of cases by 5 year age group in the ETS system in 2005, 2010 and 2015 stratified by UK birth status.", fig.cap = "Proportion of cases by 5 year age group in the ETS system in 2005, 2010 and 2015 stratified by UK birth status. Non-UK born cases have a higher proportion of young adult cases with very few cases in children or in older adults. UK born cases have a more uniform distribution of cases with some evidence of a higher proportion of cases in young adults. In the non-UK born the proportion of cases in young adults has decreased over time, with no evidence of a temporal trend in the UK born. These results are not adjusted for population demographics and therefore may be biased.", out.extra = ""}
if (regen_results) {
